# Installing Packages
```{r}
library(pageviews)
library(fpp2)
library(dplyr)
library(lubridate)
library(ggplot2)
library(forecast)
library(feasts)
```

# Getting Data
```{r}
# Monthly data
obama_pageviews_monthly <- article_pageviews(
  article = "Barack_Obama",
  granularity = "monthly",
  start = "2016010100",
  end = "2023123100"
)

## Average daily views per month
obama_ts <-ts(data = obama_pageviews_monthly$views, start = c(2016,1), frequency = 12)
obama_ts_DA <- obama_ts/monthdays(obama_ts)
obama_ts_DA
autoplot(obama_ts_DA) +
  xlab("Years") + ylab("Views") +
  ggtitle("Average views per day in a month")

# Daily data
obama_pageviews_daily <- article_pageviews(
  article = "Barack_Obama",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

## Daily timeseries
obama_ts_daily <-ts(data = obama_pageviews_daily$views, start = c(2016,01), frequency = 365.25)
autoplot(obama_ts_daily) +
  xlab("Years") + ylab("Views") +
  ggtitle("Views in a day")
```

# Preprocessing
## Splitting data
```{r}
#train-test split
obama_train_month <- window(obama_ts_DA, start=c(2016, 1), end = c(2021, 12))
obama_test_month <- window(obama_ts_DA, start = c(2022, 1), end = c(2023, 12))

#train-test split daily
obama_train_daily <- window(obama_ts_daily, start = c(2016, 1), end = c(2021, 12))
obama_test_daily <- window(obama_ts_daily, start = c(2022, 1))

```
## Decomposition using stlf on month
```{r}
autoplot(forecast(stlf(obama_train_month, method = "naive",lambda = "auto")), series = "Naive", PI = FALSE) +
autolayer(forecast(stlf(obama_train_month, method = "rwdrift",lambda = "auto")), series = "Drift", PI = FALSE) +
autolayer(forecast(stlf(obama_train_month, method = "arima",lambda = "auto")), series = "ARIMA", PI = FALSE) +
  ggtitle("Average daily views per month")
```
## Decomposition using stlf on daily
```{r}
autoplot(forecast(stlf(obama_train_daily, method = "naive",lambda = "auto")), series = "Naive", PI = FALSE) +
autolayer(forecast(stlf(obama_train_daily, method = "rwdrift",lambda = "auto")), series = "Drift", PI = FALSE) +
autolayer(forecast(stlf(obama_train_daily, method = "arima",lambda = "auto")), series = "ARIMA", PI = FALSE) +
  ggtitle("Average daily views")
```

# 4 bench mark models
## Monthly
```{r}

```
## Daily
```{r}

```

# MSTL
```{r}
autoplot(mstl(obama_train_daily))
obama_msts <- msts(obama_train_daily, seasonal.periods = c(7,12,365.25))
autoplot(obama_msts)
obama_mstl <- (mstl(msts(obama_train_daily, seasonal.periods = c(7,12,365.25))))
obama_mstl
autoplot(obama_mstl)

#splitting into seasonal and seasonally adjusted
obama_s_adj <- obama_mstl[,2] + obama_mstl[,6]
autoplot(obama_s_adj)

obama_seasonal_weekly <- obama_mstl[,3]
obama_seasonal_monthly <- obama_mstl[,4]
obama_seasonal_yearly <- obama_mstl[,5]

snaive(obama_seasonal_yearly, h = 730.5)
```



#ARIMA Model
```{r}
obama_arima <- auto.arima(obama_s_adj)

obama_s_adj_arima_fc <- forecast(obama_arima, h = 730.5)
plot(obama_s_adj_arima_fc)
obama_arima_fc <- obama_s_adj_arima_fc$mean + snaive(obama_seasonal_yearly, h = 730.5)$mean
obama_arima_fc <- ts(obama_arima_fc, start = c(2022, 1), frequency = 365.25)
#autoplot(obama_ts_daily) + 
autoplot(obama_arima_fc) + autolayer(obama_test_daily)
MAE(obama_arima_fc, obama_test_daily)
```
MAE of ARIMA model is 55327.6


#ARIMA-X Model
```{r}
# Daily data for Democratic Party pageviews
dem_pageviews_daily <- article_pageviews(
  article = "Democratic_Party_(United_States)",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

dem_ts_daily <- ts(data = dem_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

obama_arimax <- auto.arima(obama_s_adj, xreg = window(dem_ts_daily,start = c(2016, 1), end = c(2021, 12)))
forecast_obama_s_adj <- forecast(obama_arimax, xreg = dem_ts_daily, h = 730.5) 
forecast_obama <- forecast_obama_s_adj$mean + snaive(obama_seasonal_yearly, h = 730.5)
forecast_obama <- ts(forecast_obama, start = c(2022, 1), frequency = 365.25)
autoplot(forecast_obama) + autolayer(obama_test_daily)
MAE(forecast_obama, obama_test_daily)

```
MAE of ARIMA-X model is 37994.29, which is a significant improvement to ARIMA. We will be using the party pageviews as the exogenous variable to forecast the candidates' pageviews.

#Forecasting for different Presidential Candidates in 2024
List of candidates:

Democrats                      |                 Republicans
Marianne Williamson                              Donald Trump
Jason Palmer
Joe Biden
```{r}
# Daily data for Williamson
williamson_pageviews_daily <- article_pageviews(
  article = "Marianne_Williamson",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

williamson_ts_daily <- ts(data = williamson_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

# Daily data for Palmer
palmer_pageviews_daily <- article_pageviews(
  article = "Jason_Palmer",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

palmer_ts_daily <- ts(data = palmer_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

# Daily data for Joe Biden
biden_pageviews_daily <- article_pageviews(
  article = "Joe_Biden",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

biden_ts_daily <- ts(data = biden_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

# Daily data for Donald Trump
trump_pageviews_daily <- article_pageviews(
  article = "Donald_Trump",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

trump_ts_daily <- ts(data = biden_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

# Daily data for Republican Party
trump_pageviews_daily <- article_pageviews(
  article = "Republican_Party_(United_States)",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

trump_ts_daily <- ts(data = biden_pageviews_daily$views, start = c(2016,1), frequency = 365.25)
```


