# Installing Packages
```{r}
library(pageviews)
library(fpp2)
library(dplyr)
library(lubridate)
library(ggplot2)
library(forecast)
library(feasts)
library(tsibble)
```

# Getting Data
```{r}
# Monthly data
obama_pageviews_monthly <- article_pageviews(
  article = "Barack_Obama",
  granularity = "monthly",
  start = "2016010100",
  end = "2023123100"
)

## Average daily views per month
obama_ts <-ts(data = obama_pageviews_monthly$views, start = c(2016,1), frequency = 12)
obama_ts_DA <- obama_ts/monthdays(obama_ts)
obama_ts_DA
autoplot(obama_ts_DA) +
  xlab("Years") + ylab("Views") +
  ggtitle("Average views per day in a month")

# Daily data
obama_pageviews_daily <- article_pageviews(
  article = "Barack_Obama",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

## Daily timeseries
obama_ts_daily <-ts(data = obama_pageviews_daily$views, start = c(2016,01), frequency = 365.25)
autoplot(obama_ts_daily) +
  xlab("Years") + ylab("Views") +
  ggtitle("Views in a day")

obama_pageviews_daily$date[which.max(obama_pageviews_daily$views)]
```

# Preprocessing
## Splitting data
```{r}
#train-test split
obama_train_month <- window(obama_ts_DA, start=c(2016, 1), end = c(2022, 0))
obama_train_month
obama_test_month <- window(obama_ts_DA, start = c(2022, 1))

#train-test split daily
obama_train_daily <- window(obama_ts_daily, start = c(2016, 1), end = c(2022, 0))
obama_test_daily <- window(obama_ts_daily, start = c(2022, 1))

```
## Decomposition using stlf on month
```{r}
autoplot(obama_train_month) +
autolayer(forecast(stlf(obama_train_month, method = "naive",lambda = "auto"), h =24), series = "Naive", PI = FALSE) +
autolayer(forecast(stlf(obama_train_month, method = "rwdrift",lambda = "auto"), h =24), series = "Drift", PI = FALSE) +
autolayer(forecast(stlf(obama_train_month, method = "arima",lambda = "auto"), h =24), series = "ARIMA", PI = FALSE) +
  ggtitle("Average daily views per month")
```
## Decomposition using stlf on daily
```{r}
autoplot(obama_train_daily) +
autolayer(forecast(stlf(obama_train_daily, method = "naive",lambda = "auto"), h = 730), series = "Naive", PI = FALSE) +
autolayer(forecast(stlf(obama_train_daily, method = "rwdrift",lambda = "auto"), h = 730), series = "Drift", PI = FALSE) +
autolayer(forecast(stlf(obama_train_daily, method = "arima",lambda = "auto"), h = 730), series = "ARIMA", PI = FALSE) +
  ggtitle("Average daily views")
```

# 4 bench mark models
## Monthly
```{r}
obama_month_fitted_n <- forecast(stlf(obama_train_month, method = "naive",lambda = "auto"), h =24)$mean
obama_month_fitted_d <- forecast(stlf(obama_train_month, method = "rwdrift",lambda = "auto"), h =24)$mean
obama_month_fitted_a <-forecast(stlf(obama_train_month, method = "arima",lambda = "auto"), h =24)$mean

# Accuracy computation
accuracy(obama_month_fitted_n, obama_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(obama_month_fitted_d, obama_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(obama_month_fitted_a, obama_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
# Negative ME from ARIMA model suggests that it tends to underestimate the quantity being predicted. However naive seems to give the lowest MAPE of 23.164.

```
## Daily
```{r}
obama_daily_fitted_n <- forecast(stlf(obama_train_daily, method = "naive",lambda = "auto"), h = 730)$mean
obama_daily_fitted_d <- forecast(stlf(obama_train_daily, method = "rwdrift",lambda = "auto"), h = 730)$mean
obama_daily_fitted_a <-forecast(stlf(obama_train_daily, method = "arima",lambda = "auto"), h = 730)$mean

# Accuracy computation
accuracy(obama_daily_fitted_n, obama_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(obama_daily_fitted_d, obama_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(obama_daily_fitted_a, obama_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
# Negative ME from ARIMA model suggests that it tends to underestimate the quantity being predicted, RWF has the smallest MAPE of 24.658
```

# Splines
```{r}
# The daily time series has too many data points, hence too long to compute training set fits and residuals. Moving forward only monthly daily average will be computed.
obama_month_fitted_s <- splinef(obama_train_month, h = 24,lambda = "auto")$mean
obama_month_fitted_ns <- obama_month_fitted_n +obama_month_fitted_s
obama_month_fitted_as <- obama_month_fitted_a +obama_month_fitted_s
obama_month_fitted_ds <-obama_month_fitted_d +obama_month_fitted_s
autoplot(obama_test_month) +
  autolayer(obama_month_fitted_s) +
  autolayer(obama_month_fitted_ns) +
  autolayer(obama_month_fitted_as) +
  autolayer(obama_month_fitted_ds) +
  autolayer(obama_month_fitted_a)



accuracy(obama_month_fitted_s, obama_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(obama_month_fitted_ns, obama_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(obama_month_fitted_as, obama_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(obama_month_fitted_ds, obama_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
#Lowest MAPE is surprisingly naive + spline with 31.46659 though when eyeballing the ARIMA + spline looks closer to the data.

# Hence for prediction of pageviews on a daily average, a standard ARIMA model seems to perform better.

```


# MSTL
```{r}
autoplot(mstl(obama_train_daily))
obama_msts <- msts(obama_train_daily, seasonal.periods = c(7,12,365.25), start = c(2016,1))
autoplot(obama_msts)
obama_mstl <- mstl(obama_msts)

obama_mstl
autoplot(obama_mstl)

#splitting into seasonal and seasonally adjusted
obama_s_adj <- obama_mstl[,2] + obama_mstl[,6]
autoplot(obama_s_adj)

obama_seasonal_weekly <- obama_mstl[,3]
obama_seasonal_monthly <- obama_mstl[,4]
obama_seasonal_yearly <- obama_mstl[,5]

forecast_weekly <- ts(snaive(obama_seasonal_weekly, h = 730.5)$mean, start=c(2022, 1), frequency = 365.25)
forecast_monthly <- ts(snaive(obama_seasonal_monthly, h = 730.5)$mean, start=c(2022, 1), frequency = 365.25)
forecast_yearly <- ts(snaive(obama_seasonal_yearly, h = 730.5)$mean, start=c(2022, 1), frequency = 365.25)

obama_arima <- auto.arima(obama_s_adj)

forecast_s_adj <- ts((forecast(obama_arima, h = 730.5))$mean, start=c(2022, 1), frequency = 365.25)
autoplot(obama_train_daily) +
  autolayer(forecast_weekly, series = "weekly") +
  autolayer(forecast_monthly, series = "monthly") +
  autolayer(forecast_yearly, series = "yearly") +
  autolayer(forecast_s_adj, series = "Seasonal adj")

obama_mstl_fitted <- forecast_weekly + forecast_monthly + forecast_yearly + forecast_s_adj

autoplot(obama_mstl_fitted) +
  autolayer(obama_test_daily)

accuracy(obama_mstl_fitted,obama_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
#MAPE of 170.0686

autoplot(obama_seasonal_monthly)
autoplot(obama_seasonal_weekly)
autoplot(obama_seasonal_yearly)
```


#ARIMA Model
```{r}
obama_arima <- auto.arima(obama_s_adj)

obama_s_adj_arima_fc <- (forecast(obama_arima, h = 730.5))
plot(obama_s_adj_arima_fc)
obama_arima_fc <- obama_s_adj_arima_fc$mean + snaive(obama_seasonal_yearly, h = 730.5)$mean
obama_arima_fc <- ts(obama_arima_fc, start = c(2022, 1), frequency = 365.25)
autoplot(obama_arima_fc) + autolayer(obama_test_daily)
accuracy(obama_arima_fc, obama_test_daily)
```
MAPE of ARIMA model is 157.8397


#ARIMA-X Model
```{r}
# Daily data for Democratic Party pageviews
dem_pageviews_daily <- article_pageviews(
  article = "Democratic_Party_(United_States)",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)
obama_s_adj
dem_ts_daily <- ts(data = dem_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

obama_arimax <- auto.arima(obama_s_adj, xreg = window(dem_ts_daily,start = c(2016, 1), end = c(2022, 0)))
forecast_obama_s_adj <- forecast(obama_arimax, xreg = dem_ts_daily, h = 730.5) 
forecast_obama <- forecast_obama_s_adj$mean + snaive(obama_seasonal_yearly, h = 730.5)$mean 
forecast_obama <- pmax(ts(forecast_obama, start = c(2022, 1), frequency = 365.25), 0)
autoplot(obama_test_daily) + autolayer(forecast_obama)
forecast_obama
accuracy(forecast_obama, obama_test_daily)

```
MAPE of ARIMA-X model is 70.45734, which is a significant improvement to ARIMA. We will be using the party pageviews as the exogenous variable to forecast the candidates' pageviews.


#Forecasting for different Presidential Candidates in 2024
List of candidates:

Democrats                      |                 Republicans
Marianne Williamson                              Donald Trump
Jason Palmer
Joe Biden

## 1. Williamson
```{r}
# Daily data for Williamson
williamson_pageviews_daily <- article_pageviews(
  article = "Marianne_Williamson",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

williamson_ts_daily <- ts(data = williamson_pageviews_daily$views, start = c(2016,1), frequency = 365.25)
williamson_train_daily <- window(williamson_ts_daily, start = c(2016, 1), end = c(2022, 0))
williamson_test_daily <- window(williamson_ts_daily, start = c(2022, 1))

# Monthly data
williamson_pageviews_monthly <- article_pageviews(
  article = "Marianne_Williamson",
  granularity = "monthly",
  start = "2016010100",
  end = "2023123100"
)

## Average daily views per month
williamson_ts <-ts(data = williamson_pageviews_monthly$views, start = c(2016,1), frequency = 12)
williamson_ts_DA <- williamson_ts/monthdays(williamson_ts)
williamson_ts_DA
autoplot(williamson_ts_DA) +
  xlab("Years") + ylab("Views") +
  ggtitle("Average views per day in a month")

# Monthly
williamson_train_month <- window(williamson_ts_DA, start = c(2016, 1), end = c(2021, 12))
williamson_test_month <- window(williamson_ts_DA, start = c(2022, 1))

williamson_ts_daily <- ts(data = williamson_pageviews_daily$views, start = c(2016,1), frequency = 365.25)
```


### Benchmark Models
#### Monthly
```{r}
williamson_month_fitted_n <- forecast(stlf(williamson_train_month, method = "naive",lambda = "auto"), h =24)$mean
williamson_month_fitted_d <- forecast(stlf(williamson_train_month, method = "rwdrift",lambda = "auto"), h =24)$mean
williamson_month_fitted_a <-forecast(stlf(williamson_train_month, method = "arima",lambda = "auto"), h =24)$mean

# Accuracy computation
accuracy(williamson_month_fitted_n, williamson_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(williamson_month_fitted_d, williamson_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(williamson_month_fitted_a, williamson_test_month)[1,c("ME", "RMSE","MAE","MAPE")]

# Drift has the smallest MAPE of 51.24148
```


#### Daily
```{r}
williamson_daily_fitted_n <- forecast(stlf(williamson_train_daily, method = "naive",lambda = "auto"), h =730)$mean
williamson_daily_fitted_d <- forecast(stlf(williamson_train_daily, method = "rwdrift",lambda = "auto"), h =730)$mean
williamson_daily_fitted_a <-forecast(stlf(williamson_train_daily, method = "arima",lambda = "auto"), h =730)$mean

autoplot(williamson_daily_fitted_n)
# Accuracy computation
accuracy(williamson_daily_fitted_n, williamson_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(williamson_daily_fitted_d, williamson_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(williamson_daily_fitted_a, williamson_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]

# ARIMA has the smallest MAPE of 53.14829. 
```

### MSTL
```{r}
williamson_msts <- msts(williamson_train_daily, seasonal.periods = c(7,12,365.25), start = c(2016,01))
williamson_mstl <- mstl(williamson_msts)

autoplot(williamson_mstl)
#splitting into seasonal and seasonally adjusted
williamson_s_adj <- williamson_mstl[,2] + williamson_mstl[,6]

williamson_seasonal_weekly <- williamson_mstl[,3]
williamson_seasonal_monthly <- williamson_mstl[,4]
williamson_seasonal_yearly <- williamson_mstl[,5]

forecast_weekly_w <- snaive(williamson_seasonal_weekly , h = 730.5)$mean
forecast_monthly_w <- snaive(williamson_seasonal_monthly, h = 730.5)$mean
forecast_yearly_w <- snaive(williamson_seasonal_yearly, h = 730.5)$mean

williamson_arima <- auto.arima(williamson_s_adj)

forecast_s_adj_w <- (forecast(williamson_arima, h = 730.5))$mean
autoplot(williamson_train_daily) +
  autolayer(forecast_weekly_w, series = "weekly") +
  autolayer(forecast_monthly_w, series = "monthly") +
  autolayer(forecast_yearly_w, series = "yearly") +
  autolayer(forecast(williamson_arima, h = 730.5), series = "Seasonal adj")

williamson_mstl_fitted <- forecast_weekly_w + forecast_monthly_w + forecast_yearly_w + forecast_s_adj_w

autoplot(williamson_mstl_fitted) +
  autolayer(williamson_test_daily)

accuracy(williamson_mstl_fitted,williamson_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
# MAPE of 161.4983.
```

### ARIMA-X
```{r}
williamson_arimax <- auto.arima(williamson_s_adj, xreg = window(dem_ts_daily,start = c(2016, 1), end = c(2022, 0)))
forecast_williamson_s_adj <- forecast(williamson_arimax, xreg = dem_ts_daily, h = 730.5) 
forecast_williamson <- forecast_williamson_s_adj$mean + snaive(williamson_seasonal_yearly, h = 730.5)$mean
forecast_williamson <- pmax(ts(forecast_williamson, start = c(2022, 1), frequency = 365.25), 0)
autoplot(williamson_test_daily) + autolayer(forecast_williamson)
accuracy(forecast_williamson, williamson_test_daily)


williamson_full <- (mstl(msts(williamson_ts_daily, seasonal.periods = c(7,12,365.25))))
williamson_full_s_adj <- williamson_full[,2] + williamson_full[,6]
williamson_seasonal_yearly_full <- williamson_full[,5]
williamson_arimax_full <- auto.arima(williamson_full_s_adj, xreg = dem_ts_daily)
forecast_williamson_s_adj_full <- forecast(williamson_arimax_full, xreg = dem_ts_daily, h = 365.25) 
forecast_williamson_full <- forecast_williamson_s_adj_full$mean + snaive(williamson_seasonal_yearly_full, h = 365.25)$mean
forecast_williamson_full <- pmax(ts(forecast_williamson_full, start = c(2024, 1), frequency = 365.25), 0)
autoplot(forecast_williamson_full)
#MAPE of Williamson's ARIMA-X is 126.5206
```


## 2. Jason Palmer <--- Wikipedia page was created 6 March 2024. We will not be able to forecast for Palmer.

```{r}
# Daily data for Palmer
palmer_pageviews_daily <- article_pageviews(
  article = "Jason_Palmer_(politician)",
  granularity = "daily",
  start = "2024010100",
  end = "2024123100"
)

palmer_ts_daily <- ts(data = palmer_pageviews_daily$views, start = c(2024,1), frequency = 365.25)
palmer_ts_daily

```

## 3. Joe Biden
```{r}
# Daily data for Joe Biden
biden_pageviews_daily <- article_pageviews(
  article = "Joe_Biden",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

biden_ts_daily <- ts(data = biden_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

biden_train_daily <- window(biden_ts_daily, start = c(2016, 1), end = c(2022, 0))
autoplot(biden_train_daily)
biden_test_daily <- window(biden_ts_daily, start = c(2022, 1))

biden_msts <- msts(biden_train_daily, seasonal.periods = c(7,12,365.25))
biden_mstl <- (mstl(msts(biden_train_daily, seasonal.periods = c(7,12,365.25))))

# Monthly data
biden_pageviews_monthly <- article_pageviews(
  article = "Joe_Biden",
  granularity = "monthly",
  start = "2016010100",
  end = "2023123100"
)

## Average daily views per month
biden_ts <-ts(data = biden_pageviews_monthly$views, start = c(2016,1), frequency = 12)
biden_ts_DA <- biden_ts/monthdays(biden_ts)
biden_ts_DA
autoplot(biden_ts_DA) +
  xlab("Years") + ylab("Views") +
  ggtitle("Average views per day in a month")

# Monthly
biden_train_month <- window(biden_ts_DA, start = c(2016, 1), end = c(2021, 12))
biden_test_month <- window(biden_ts_DA, start = c(2022, 1))

biden_ts_daily <- ts(data = biden_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

```

### Benchmark Models
#### Monthly
```{r}
biden_month_fitted_n <- forecast(stlf(biden_train_month, method = "naive",lambda = "auto"), h =24)$mean
biden_month_fitted_d <- forecast(stlf(biden_train_month, method = "rwdrift",lambda = "auto"), h =24)$mean
biden_month_fitted_a <-forecast(stlf(biden_train_month, method = "arima",lambda = "auto"), h =24)$mean

# Accuracy computation
accuracy(biden_month_fitted_n, biden_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(biden_month_fitted_d, biden_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(biden_month_fitted_a, biden_test_month)[1,c("ME", "RMSE","MAE","MAPE")]

# Drift has the smallest MAPE of 24.42426.
```

#### Daily
```{r}
biden_daily_fitted_n <- forecast(stlf(biden_train_daily, method = "naive",lambda = "auto"), h =730)$mean
biden_daily_fitted_d <- forecast(stlf(biden_train_daily, method = "rwdrift",lambda = "auto"), h =730)$mean
biden_daily_fitted_a <-forecast(stlf(biden_train_daily, method = "arima",lambda = "auto"), h =730)$mean

# Accuracy computation
accuracy(biden_daily_fitted_n, biden_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(biden_daily_fitted_d, biden_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(biden_daily_fitted_a, biden_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]

# ARIMA has the smallest MAPE of 29.42795
```

### MSTL with ARIMA
```{r}
biden_msts <- msts(biden_train_daily, seasonal.periods = c(7,12,365.25), start = c(2016,1))
biden_mstl <- mstl(biden_msts)

autoplot(biden_mstl)
#splitting into seasonal and seasonally adjusted
biden_s_adj <- biden_mstl[,2] + biden_mstl[,6]

biden_seasonal_weekly <- biden_mstl[,3]
biden_seasonal_monthly <- biden_mstl[,4]
biden_seasonal_yearly <- biden_mstl[,5]

forecast_weekly_b <- ts(snaive(biden_seasonal_weekly , h = 730.5)$mean, start = c(2022, 1), frequency = 365.25)
forecast_monthly_b <- ts(snaive(biden_seasonal_monthly, h = 730.5)$mean, start = c(2022, 1), frequency = 365.25)
forecast_yearly_b <- ts(snaive(biden_seasonal_yearly, h = 730.5)$mean, start = c(2022, 1), frequency = 365.25)

biden_arima <- auto.arima(biden_s_adj)

forecast_s_adj_b <- ts((forecast(biden_arima, h = 730.5))$mean, start = c(2022, 1), frequency = 365.25)
autoplot(biden_train_daily) +
  autolayer(forecast_weekly_b, series = "weekly") +
  autolayer(forecast_monthly_b, series = "monthly") +
  autolayer(forecast_yearly_b, series = "yearly") +
  autolayer(forecast_s_adj_b, series = "Seasonal adj")

biden_mstl_fitted <- forecast_weekly_b + forecast_monthly_b + forecast_yearly_b + forecast_s_adj_b
biden_mstl_fitted <- ts(biden_mstl_fitted, start = c(2022, 1), frequency = 365.25)

autoplot(biden_test_daily) + autolayer(biden_mstl_fitted)


accuracy(biden_mstl_fitted,biden_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
# MAPE of 155.4375.
```

#ARIMA-X
```{r}
biden_arimax <- auto.arima(biden_s_adj, xreg = window(dem_ts_daily,start = c(2016, 1), end = c(2022, 0)))
forecast_biden_s_adj <- forecast(biden_arimax, xreg = dem_ts_daily, h = 730.5) 
forecast_biden <- forecast_biden_s_adj$mean + snaive(biden_seasonal_yearly, h = 730.5)$mean
forecast_biden <- pmax(ts(forecast_biden, start = c(2022, 1), frequency = 365.25), 0)
forecast_biden
accuracy(forecast_biden, biden_test_daily)
#MAPE of 89.12801

biden_full <- (mstl(msts(biden_ts_daily, seasonal.periods = c(7,12,365.25))))
biden_full_s_adj <- biden_full[,2] + biden_full[,6]
biden_seasonal_yearly_full <- biden_full[,5]
biden_arimax_full <- auto.arima(biden_full_s_adj, xreg = dem_ts_daily)
forecast_biden_s_adj_full <- forecast(biden_arimax_full, xreg = dem_ts_daily, h = 365.25) 
forecast_biden_full <- forecast_biden_s_adj_full$mean + snaive(biden_seasonal_yearly_full, h = 365.25)$mean
forecast_biden_full <- pmax(ts(forecast_biden_full, start = c(2024, 1), frequency = 365.25), 0)
autoplot(biden_ts_daily) + autolayer(forecast_biden_full)
```

## 4. Donald Trump
```{r}
# Daily data for Republican Party
rep_pageviews_daily <- article_pageviews(
  article = "Republican_Party_(United_States)",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

rep_ts_daily <- ts(data = rep_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

# Daily data for Donald Trump
trump_pageviews_daily <- article_pageviews(
  article = "Donald_Trump",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

trump_ts_daily <- ts(data = trump_pageviews_daily$views, start = c(2016,1), frequency = 365.25)

trump_train_daily <- window(trump_ts_daily, start = c(2016, 1), end = c(2022, 0))
trump_test_daily <- window(trump_ts_daily, start = c(2022, 1))
# Monthly data
trump_pageview_monthly <- article_pageviews(
  article = "Donald_Trump",
  granularity = "monthly",
  start = "2016010100",
  end = "2023123100"
)

## Average daily views per month
trump_ts <-ts(data = trump_pageview_monthly$views, start = c(2016,1), frequency = 12)
trump_ts_DA <- trump_ts/monthdays(trump_ts)
trump_ts_DA
autoplot(trump_ts_DA) +
  xlab("Years") + ylab("Views") +
  ggtitle("Average views per day in a month")

# Monthly
trump_train_month <- window(trump_ts_DA, start = c(2016, 1), end = c(2021, 12))
trump_test_month <- window(trump_ts_DA, start = c(2022, 1))

```
### Benchmark Models
#### Monthly
```{r}
trump_month_fitted_n <- forecast(stlf(trump_train_month, method = "naive",lambda = "auto"), h =24)$mean
trump_month_fitted_d <- forecast(stlf(trump_train_month, method = "rwdrift",lambda = "auto"), h =24)$mean
trump_month_fitted_a <-forecast(stlf(trump_train_month, method = "arima",lambda = "auto"), h =24)$mean

# Accuracy computation
accuracy(trump_month_fitted_n, trump_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(trump_month_fitted_d, trump_test_month)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(trump_month_fitted_a, trump_test_month)[1,c("ME", "RMSE","MAE","MAPE")]

#Naive has the smallest MAPE of 180.7104. 
```

#### Daily
```{r}
trump_daily_fitted_n <- forecast(stlf(trump_train_daily, method = "naive",lambda = "auto"), h =730)$mean
trump_daily_fitted_d <- forecast(stlf(trump_train_daily, method = "rwdrift",lambda = "auto"), h =730)$mean
trump_daily_fitted_a <-forecast(stlf(trump_train_daily, method = "arima",lambda = "auto"), h =730)$mean

# Accuracy computation
accuracy(trump_daily_fitted_n, trump_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(trump_daily_fitted_d, trump_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
accuracy(trump_daily_fitted_a, trump_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]

#ARIMA has the smallest MAPE of 194.6183
```
### MSTL
```{r}
trump_msts <- msts(trump_train_daily, seasonal.periods = c(7,12,365.25), start = c(2016,01))
trump_mstl <- mstl(trump_msts)

autoplot(trump_mstl)
#splitting into seasonal and seasonally adjusted
trump_s_adj <- trump_mstl[,2] + trump_mstl[,6]

trump_seasonal_weekly <- trump_mstl[,3]
trump_seasonal_monthly <- trump_mstl[,4]
trump_seasonal_yearly <- trump_mstl[,5]

forecast_weekly_t <- snaive(trump_seasonal_weekly , h = 730.5)$mean
forecast_monthly_t <- snaive(trump_seasonal_monthly, h = 730.5)$mean
forecast_yearly_t <- snaive(trump_seasonal_yearly, h = 730.5)$mean

trump_arima <- auto.arima(trump_s_adj)

forecast_s_adj_t <- (forecast(trump_arima, h = 730.5))$mean
autoplot(trump_train_daily) +
  autolayer(forecast_weekly_t, series = "weekly") +
  autolayer(forecast_monthly_t, series = "monthly") +
  autolayer(forecast_yearly_t, series = "yearly") +
  autolayer(forecast(trump_arima, h = 730.5), series = "Seasonal adj")

trump_mstl_fitted <- forecast_weekly_t + forecast_monthly_t + forecast_yearly_t + forecast_s_adj_t

autoplot(trump_mstl_fitted) +
  autolayer(trump_test_daily)

accuracy(trump_mstl_fitted,trump_test_daily)[1,c("ME", "RMSE","MAE","MAPE")]
#MAPE of 204.5591 .
```

#ARIMA-X
```{r}

trump_arimax <- auto.arima(trump_s_adj, xreg = window(rep_ts_daily,start = c(2016, 1), end = c(2022, 0)))
forecast_trump_s_adj <- forecast(trump_arimax, xreg = rep_ts_daily, h = 730.5) 
forecast_trump <- forecast_trump_s_adj$mean + snaive(trump_seasonal_yearly, h = 730.5)$mean
forecast_trump <- pmax(ts(forecast_trump, start = c(2022, 1), frequency = 365.25), 0)
forecast_trump
accuracy(forecast_trump, trump_test_daily)
#MAPE of 183.5721

trump_full <- (mstl(msts(trump_ts_daily, seasonal.periods = c(7,12,365.25))))
trump_full_s_adj <- trump_full[,2] + trump_full[,6]
trump_seasonal_yearly_full <- trump_full[,5]
trump_arimax_full <- auto.arima(trump_full_s_adj, xreg = rep_ts_daily)
forecast_trump_s_adj_full <- forecast(trump_arimax_full, xreg = rep_ts_daily, h = 365.25) 
forecast_trump_full <- forecast_trump_s_adj_full$mean + snaive(trump_seasonal_yearly_full, h = 365.25)$mean
forecast_trump_full <- pmax(ts(forecast_trump_full, start = c(2024, 1), frequency = 365.25), 0)
autoplot(forecast_trump_full) + autolayer(trump_ts_daily)


```

#Data for Hillary Clinton
```{r}
# Daily data for Clinton
clinton_pageviews_daily <- article_pageviews(
  article = "Hillary Clinton",
  granularity = "daily",
  start = "2016010100",
  end = "2023123100"
)

clinton_ts_daily <- ts(data = clinton_pageviews_daily$views, start = c(2016,1), frequency = 365.25)
autoplot(window(clinton_ts_daily, start = c(2016, 1), end = c(2017, 1))) + autolayer(window(trump_ts_daily, start = c(2016, 1), end = c(2017, 1)))
```















