<<<<<<< Updated upstream
# Installing Packages
```{r}
# library(httr)
# library(jsonlite)
library(pageviews)
library(fpp2)
library(dplyr)
library(lubridate)
library(ggplot2)
library(forecast)
```
# Project is about US presidential election, looking at page views and trying to predict web traffic

2008
Barack Obama (Democrat, Winner) and Joe Biden (Running Mate)
John McCain (Republican) and Sarah Palin (Running Mate)

2012
Barack Obama (Democrat, Winner) and Joe Biden (Running Mate)
Mitt Romney (Republican) and Paul Ryan (Running Mate)


2016
Donald Trump (Republican, Winner) and Mike Pence (Running Mate)
Hillary Clinton (Democrat) and Tim Kaine (Running Mate)

-- Due to lack of credible historic data we will not use oldpage views and hence not forecast 2008, 2012, 2016 -- 

2020
Joe Biden (Democrat, Winner) and Kamala Harris (Running Mate)
Donald Trump (Republican) and Mike Pence (Running Mate)

Election-Related Pages
In addition to the candidate pages, consider analyzing traffic to pages related to the elections themselves and key issues that were prominent during those times:

Pages for each election (e.g., "2008 United States presidential election", "2012 United States presidential election", etc.).
Key political and election issues (e.g., "Healthcare reform in the United States", "United States foreign policy", "Climate change policy of the United States", "Immigration policy of Donald Trump", etc.) during the election years.
Political party pages (e.g., "Democratic Party (United States)", "Republican Party (United States)").


```{r}
# old_pageviews looks at data from January 2008 to July 2016
#Description
# This retrieves per-project pageview counts from January 2008 to July 2016. These counts are calculated using the 'legacy' (read: old) model, which overcounts due to its inclusion of web-crawlers and similar automata.


pageview_timestamps(timestamps = Sys.Date(), first = TRUE)
# Basic example
r_pageviews <- article_pageviews()
# 2015-07-01 is the earliest it can access

```
Steps
1. Get pageview data --> convert to time series --> daily average --> eyeball data --> determine type of decomposition
2. Make data stationary -->nsdiffs, ndiffs --> makes ts variance stationary
3. BoxCox Transformation --> keep variance constant (for ARIMA)
4. Decomposition --> if amplitude constant over time --> additive; if increases or decreases --> multiplicative --> split into seasonal + seasonal adj
5. Perform Ljung Box test on residuals --> (observed - fitted both seasonal and seasonal adj) --> if fail
3. 
4. decompose ts --> split into seasonal and seasonal_adj
5. run Ljung_box on residuals (observed - fitted) --> if fail --> move on  to additive --> if fail move on to other models

Steps (ARIMA)
1. Get pageview data --> convert to time series --> daily average --> eyeball data
2. Determine if box.cox transformation is needed --> make variance stationary
3. Either AUTO ARIMA or manual
4. AUTO ARIMA --> determine model
ARIMA
5. nsdiffs --> determine if ARIMA or SARIMA model (D component, D = 0 then ARIMA)
6. ndiffs --> determine order of integration (d component)
7. ACF and PACF --> determines AR component (p, q component find significant lags, determine if exponentially decaying or spike at significant lags or smooth decaying)
8. Model Selection --> look at slides first
# Page Views from January 2008 to July 2016 -- Democratic
```{r}
# 2020
## Democrats

# Obama announced that he would run for president in Feb 2007

obama_pageviews <- article_pageviews(
  article = "Barack_Obama",
  granularity = "monthly",
  start = "2016010100",
  end = "2023123100"
)

tail(obama_pageviews)

obama_ts <-ts(data = obama_pageviews$views, start = c(2016,1), frequency = 12)
obama_ts_DA <- obama_ts/monthdays(obama_ts)
obama_ts_DA
autoplot(obama_ts_DA) +
  xlab("Years") + ylab("Views") +
  ggtitle("Average views per day")

# There are significant lags in r1,2,3 and in r19 to 24
obama_lambda <- BoxCox.lambda(obama_ts_DA)
obama_bc <- BoxCox(obama_ts_DA, obama_lambda)
autoplot(obama_bc)

```

## Preprocessing
```{r}
#train-test split
obama_train <- window(obama_bc, start=c(2016, 1), end = c(2021, 12))
obama_test <- window(obama_ts_DA, start = c(2022, 1), end = c(2023, 12))

```

# Build the 4 benchmark models

```{r}
obama_train_mean <- meanf(obama_train, h = 24)
obama_train_naive <- naive(obama_train, h = 24)
obama_train_snaive <- snaive(obama_train, h = 24)
obama_train_drift <- rwf(obama_train, drift = TRUE, h = 24)

autoplot(obama_train) +
  autolayer(obama_train_mean, series = "Mean", PI = FALSE) +
  autolayer(obama_train_naive, series = "Naive", PI = FALSE) +
  autolayer(obama_train_snaive, series = "Seasonal Naive", PI = FALSE) +
  autolayer(obama_train_drift, series = "Drift", PI = FALSE) +
  xlab("Year") + ylab("Average Daily Views") + 
  ggtitle("Forecasts for Average Daily Views on Barrack Obama") +
  guides(colour = guide_legend(title = "Forecast"))

## Mean
checkresiduals(meanf(obama_train, h = 24), series = "Mean", PI = FALSE) # There is significant autocorrelation, residuals are normally distributed, mean of the residuals is close to 0
accuracy(InvBoxCox(unlist(obama_train_mean$mean), lambda = obama_lambda), obama_test)

## Naive 
checkresiduals(naive(obama_train, h = 24), series = "Naive", PI = FALSE) # There is no significant autocorrelation else for the first lag,residuals are relatively normally distributed, mean of the residuals is close to 0
accuracy(InvBoxCox(unlist(obama_train_naive$mean), lambda = obama_lambda), obama_test)

## Snaive
checkresiduals(snaive(obama_train, h = 24), series = "Seasonal Naive", PI = FALSE) # There is significant autocorrelation in a decaying pattern, residuals are normally distributed, mean of the residuals is close to 0
accuracy(InvBoxCox(unlist(obama_train_snaive$mean), lambda = obama_lambda), obama_test)

## Drift
checkresiduals(rwf(obama_train, drift = TRUE, h = 24), series = "Drift", PI = FALSE) # There is significant autocorrelation for the first lag, residuals are normally distributed, mean of the residuals is close to 0
accuracy(InvBoxCox(unlist(obama_train_drift$mean), lambda = obama_lambda), obama_test)

# Naive has the lowest MAE compared to 
```

## Decomposing multiplicative
```{r}
# Decompose
autoplot(decompose(x = obama_ts_DA, type = "multiplicative")) # amplitude appears to be decreasing over time, hence multiplicative

autoplot(stl(obama_ts_DA, s.window="periodic"))

obama_seasonal_component <- seasonal(decompose(x = obama_ts_DA, type = "multiplicative"))

obama_seasonaladj_component <- seasadj(decompose(x = obama_ts_DA, type = "multiplicative"))

multiplicativefitted <- (snaive(obama_seasonal_component))$fitted * (naive(obama_seasonaladj_component))$fitted

obama_residuals <- obama_ts_DA - multiplicativefitted

autoplot(obama_ts_DA) + autolayer(multiplicativefitted)

Box.test(obama_residuals, type = c("Ljung-Box")) # p-value is 0.01328 < 0.05 reject the null hypothesis. statistically significant autocorrelation in residuals at some lag.
```
## Decomposing additive
```{r}

autoplot(decompose(x = obama_ts_DA, type = "additive")) 
obama_seasonal_component_additive <- seasonal(decompose(x = obama_ts_DA, type = "additive"))
obama_seasonaladj_component_additive <- seasadj(decompose(x = obama_ts_DA, type = "additive"))

additivefitted <- (snaive(obama_seasonal_component_additive))$fitted + (naive(obama_seasonaladj_component_additive))$fitted

obama_residuals_additive <- obama_ts_DA - additivefitted
Box.test(obama_residuals_additive, type = c("Ljung-Box")) # p-value is 0.008128 < 0.05 Reject the null, there is autocorrelation in residuals at some lag.

autoplot(obama_ts_DA) + autolayer(obama_residuals) + autolayer(obama_residuals_additive) # Multiplicative decomposition is indeed better

accuracy(additivefitted, x = obama_ts_DA)
accuracy(multiplicativefitted, x = obama_ts_DA) # Multiplicative has smaller MAE, RMSE, MAPE

```

# Decomposed Forecasing
```{r}
autoplot(stl(obama_train, s.window="periodic"))

obama_seasonal_component <- seasonal(decompose(x = obama_train, type = "multiplicative"))

# Check for stationary
nsdiffs(obama_seasonal_component)
ndiffs(diff(obama_seasonal_component, 1))

# obama_seasonal_component <- diff(obama_seasonal_component)

obama_seasonaladj_component <- seasadj(decompose(x = obama_train, type = "multiplicative"))

obama_seasonal_component_snaive <- snaive(obama_seasonal_component, h = 24)

obama_seasonaladj_component_mean <- (meanf(obama_seasonaladj_component, h = 24))
obama_seasonaladj_component_naive <- (naive(obama_seasonaladj_component, h = 24))
obama_seasonaladj_component_snaive <- (snaive(obama_seasonaladj_component, h = 24))
obama_seasonaladj_component_drift <- (rwf(obama_seasonaladj_component, h = 24, drift = TRUE))

multiplicativefitted_mean <- obama_seasonaladj_component_mean$mean * obama_seasonal_component_snaive$mean
multiplicativefitted_naive <- obama_seasonaladj_component_naive$mean * obama_seasonal_component_snaive$mean
multiplicativefitted_snaive <- obama_seasonaladj_component_snaive$mean * obama_seasonal_component_snaive$mean
multiplicativefitted_drift <- obama_seasonaladj_component_drift$mean * obama_seasonal_component_snaive$mean

obama_multiplicativefitted_mean_resid <- obama_test - InvBoxCox((multiplicativefitted_mean), lambda = obama_lambda)

obama_multiplicativefitted_naive_resid <- obama_test - InvBoxCox((multiplicativefitted_naive), lambda = obama_lambda)

obama_multiplicativefitted_snaive_resid <- obama_test - InvBoxCox((multiplicativefitted_snaive), lambda = obama_lambda)

obama_multiplicativefitted_drift_resid <- obama_test - InvBoxCox((multiplicativefitted_drift), lambda = obama_lambda)

autoplot(obama_test) +
  autolayer(InvBoxCox((multiplicativefitted_mean), lambda = obama_lambda), series = "Mean") +
  autolayer(InvBoxCox((multiplicativefitted_naive), lambda = obama_lambda), series = "Naive") +
  autolayer(InvBoxCox((multiplicativefitted_snaive), lambda = obama_lambda), series = "Seasonal Naive") +
  autolayer(InvBoxCox((multiplicativefitted_drift), lambda = obama_lambda), series = "Drift")
  

```
## MSTL?
```{r}
autoplot(mstl(obama_ts_DA))

obama_mstl <- mstl(obama_ts_DA)

trendcycle(obama_mstl)
remainder(obama_mstl)
seasadj(obama_mstl)
seasonal(obama_mstl)

seasadj(obama_mstl) - (remainder(obama_mstl) + trendcycle(obama_mstl))
```

## Building complex models
```{r}
# Build complex model
## ARIMA
nsdiffs(obama_train) #returns 0
ndiffs(obama_train) #returns 0
acf(obama_ts_DA, lag.max = 96) #slow decaying, indicating presence of AR.
pacf(obama_train, lag.max = 96) #random spikes, indicating seasonal MA component

Arima(obama_train, order = c(1,0,0), seasonal = c(0,0,1)) 
Arima(obama_train, order = c(1,0,1), seasonal = c(0,0,1)) 
Arima(obama_train, order = c(1,0,1), seasonal = c(1,0,1)) 

#ARIMA(1,0,1)(1,0,1) returned the lowest AICc of -1125.53

#auto arima
obama_auto_arima <- auto.arima(obama_train)
obama_auto_arima #auto arima returned ARIMA(1,0,0)(0,0,1) with AICc of -1127.92

#ARIMA forecasting
obama_arima <- Arima(obama_train, order = c(1,0,1), seasonal = c(1,0,1)) 
obama_auto_arima_fc <- forecast(obama_auto_arima, h = 24)
obama_arima_fc <- forecast(obama_arima, h = 24)
obama_arima_fc
checkresiduals(obama_arima_fc)
checkresiduals(obama_auto_arima_fc)

obama_auto_arima_error <-  (InvBoxCox(unlist(obama_auto_arima_fc$mean) , lambda = obama_lambda) - obama_test)
mean(obama_auto_arima_error)

obama_arima_error <-  (InvBoxCox(unlist(obama_arima_fc$mean) , lambda = obama_lambda) - obama_test)
mean(obama_arima_error)

#Auto ARIMA gave a lower absolute error

autoplot(InvBoxCox(unlist(obama_arima_fc$mean) , lambda = obama_lambda), series = "ARIMA(1,0,1)(1,0,1)") + autolayer(obama_test) + autolayer(InvBoxCox(unlist(obama_auto_arima_fc$mean) , lambda = obama_lambda), series = "ARIMA(1,0,0)(0,0,1)")
```



















